\chapter{Introduction}
This document was created as part of the project \emph{Research on Numerical Simulation of Downhole Drill String Dynamics}, a fundamentals research project at CNPC USA.  The project was proposed and kicked off in 2023.

The concept for a model review and evaluation resulted from conversations with members of the \osdc{} (\referencename~\cite{ref:pastusek2019a}).  A number of models have been released to the community's \emph{Github} site, however, it was not known exactly how the models compared in terms of features and performance.  At the same time, CNPC was both interested in studying drill string models and is a member of two industry-university research programs.  These programs may provide access to drilling dynamic models.  These models also need to be reviewed to determine features and investigated to determine availability.

Based on this it was decided to review the drilling dynamic models both on the \osdc{} and others that may be accessible through industry memberships.  Promising drill string models would then be selected for further evaluation and testing.

\section{Outline}
The models associated with drilling dynamics that were available on the \osdc{} Github site, as well as those potentially available through industry-university research programs were reviewed.  However, not all of these models were exercised.  Some of the models did not pertain directly to drill strings, lacked key features, or the source code was unavailable.

Two models that passed all the criteria were selected for an in-depth evaluation.  The first is the torsional model of Aarsnes-Shor.  The model has two versions available, a more mature Matlab version and a newer port to Python.  The second model is from ExxonMobil and is written in Python.  The theory and source code of both models was examined.

To compare the models, a set of Test Cases was developed.  The procedures, methods, and reasoning for selecting the Test Cases are documented.  The models were then run on the Test Cases and the results compared.  Finally, the conclusions and recommendations are reported.

\section{Objectives}
The primary motivation for CNPC USA was to evaluate soft string models.  However, in order to compare models, it is necessary to have a set of Test Cases to use as a benchmark.  In addition, well defined procedures should be used in order to allow a direction comparison.  Based on these requirements, the following objectives were adopted.
\begin{numberedlist}
	\item \emph{Review} all the models that were potential available to determine their features.
	\item \emph{Establish} a set Test Cases for comparing drill string models.
	\item \emph{Document} procedures and methods used in the comparisons.
	\item \emph{Evaluate} soft string models and assess their features and performance.
\end{numberedlist}

\section{Organization of this Document}
This document is organized in a manner that facilitates different use cases.  Together, the chapters provide a complete description of a process that starts from a model search and ends with a comparison between models.  In addition, some of the chapters are written as stand alone chapters that can be referenced independently.  Other chapters are naturally related to each other and are best read as a unit.

\chaptername~\ref{ch:modelreview} can be read as a stand alone chapter.  It contains short sections that describe each of the models that were examined.  While the models were reviewed in the context of the current project, the reviews can serve as a reference of available drilling dynamics models.

\chaptername~\ref{ch:testcases} can be read as a stand alone chapter.  The Test Cases are generalized and not specific to the particular models evaluated.  The intent is provide baseline Test Cases that can be used to compare any drill string models.  The Test Cases and procedures are defined completely within the chapter.

In \chaptername{}s~\ref{ch:aarnessshormodel},~\ref{ch:exxonmobilmodel}, and~\ref{ch:results} two different models are described in detail and compared on the Test Cases.  These chapters are best read as a unit.

\section{Important Notes on the Models}
It should be noted that these models are considered research models and are currently still in development.  As such, they were expected to lack some features a commercial code would have.  In addition, some issues were expected to be encountered.  Any potential issues reported in this document are done solely in the interest of providing feedback to the developers in hopes of helping to improve the models.  \important{At no time should any comments in this report be construed as criticisms of these models.}  On the contrary, this collaboration is viewed as demonstrating the benefits of open-source software.  A wider audience gets to make use of the software and access to the source code means issues can examined, tested, and feedback provided to the developers.

%While these models may not contain some of the features of commercial code, they capture

%The team is eternally grateful for allowing access to these models.
\section{Nomenclature}
\subsection{Steady-State and Transient Phases}
In vibration analysis, the terms \emph{transient} and \emph{steady-state} refer to the way in which the vibration of a system behaves over time.  Transient vibration is a vibration that changes over time, while steady-state vibration is a vibration that remains constant over time.

In classical mechanics and dynamics, these are well defined and understood.  Steady-state vibration refer to harmonic vibration that represents the particular part of the solution to the governing equations of motion.  The transient response is assumed to be damped out so that it has diminished to zero.

A discrete system with repeating, but non-harmonic loading, would not strictly fall into this definition.  It differs from the classical system in the steady-state phase.  The classical steady-state displays an exactly repeating pattern of behavior.  The discrete, non-harmonic system has a phase where the initial, non-repeating load has been damped out.  However, the remaining ``steady-state'' may exhibit slight differences from cycle to cycle of vibration.  \todo{Add a figure.}

However, there are obvious similarities between the classical and discrete, non-harmonic loading cases.  A clear response to machine start up and other changes in loading is evident.  Similarly, it can be seen that, given time, the systems will damp out their response to these changes. So while the classical and the discrete, non-harmonically loaded systems exhibit somewhat different traits, they are comparable in the broad sense.  As such, it is useful to speak of the two systems in the a similar manner.  \emph{Therefore, in direct analogy to classical harmonic vibration, we will refer to the transient and steady-state phases of the discrete, non-harmonic system.}

\subsection{Damping}
\label{sec:damping}
The governing equations of motion can be represented as
\begin{align}
    m\ddot{x}      + c_a\dot{x}      + k_ax      + f_f    & = f_e    \label{eq:axial_motion_GE} \\
    J\ddot{\theta} + c_t\dot{\theta} + k_t\theta + \tau_f & = \tau_e \label{eq:torsional_motion_GE}
\end{align}
where $m$ is mass, $c_a$ is axial damping coefficient, $k_a$ is axial stiffness, $f_f$ is frictional force, $f_e$ are external forces and $x$ is axial displacement, $J$ is mass moment of inertia, $c_t$ is torsional damping coefficient, $k_t$ is torsional stiffness, $\tau_f$ is frictional torque, $\tau_e$ is external torque and $\theta$ is angular displacement.  \equationname~\ref{eq:axial_motion_GE} is the equation for linear motion and \equationname~\ref{eq:torsional_motion_GE} is the analogous equation for rotational motion.

In this project, frictional force, $f_f$, and frictional torque, $\tau_f$, are defined as a force resisting the relative motion of elements sliding against each other.  It should be noted that this is the only constraint on the definition.  For example, the frictional force may or may not have a rate dependent component.  For drill string dynamics, the primarily source of friction is from drill string and wellbore contact.

%This type of friction is a function of the friction factor and normal force at the point of contact.

The terms $c_a\dot{x}$ and $c_t\dot{\theta}$, are referred to as viscous damping forces.  For drill string dynamics, this is primarily drag that results from drilling mud and cuttings.

Both friction and viscous damping are types of damping, which is a more general term referring to forces that remove energy from the system.

These definitions are provided to clarify how these terms are used \emph{in this document}.  Other sources may define them differently, use different terms altogether, or use a different name for the same concepts.

\subsection{Models}
\subsubsection{Static and Dynamic}
Models can be classified as either static or dynamic.  Static models do not account for acceleration (\staticforcebalance{}).  Dynamic models account for inertia (\dynamicforcebalance{}).  Within each of these categories, there are a variety of types of models (single body/multibody, rigid body/flexible, et cetera).  Particularly for dynamics models, there is a broad range of fidelities (or complexity).  The authors did not feel the compulsion to name or classify each of these types of dynamics models.  Rather, a simpler, more straightforward approach is taken and the features of each model reviewed are described.

\subsubsection{Steady-State and Transient Models}
``Steady-state model'' is a synonym for ``static model.'' The phrase ``transient model'' is sometimes used to mean a dynamic model that can capture transient phases.  The authors prefer the terms ``static'' and ``dynamic'' models.  Partially because they are more approachable and easier to understand.  But also as a way to prevent confusion.  The terms ``steady-state'' and ``transient'' will be reserved for the steady-state and transient \emph{phases} of a dynamic response.

\subsection{Review and Evaluation}
In this document we will use the term ``review'' in context of a model to mean a literature review and assessment of the features of the model.  The term ``evaluation'' will be used to mean the model was executed, run, tested on standardized cases, and compared to other models.

\section{Guidance and Technical Committee}
Several members of the industry and academia played key roles in this project.  Their support and guidance is greatly appreciated.  Without their assistance, this project would not have been possible.  Brief biographies are presented below.

\begin{committeemember}{Rajat Dixit}{MS from Indian Institute of Technology}{Wells Engineer (Drilling) at ExxonMobil}
Expertise includes fluids and thermal sciences, FEA, computational sciences, solid mechanics and heat transfer.
\end{committeemember}

\begin{committeemember}{Eduardo Gildin}{Ph.D. from University of Texas}{Professor at Texas A\&M}
Research interests include model reduction of large-scale dynamical systems, control and optimization of large-scale dynamical systems and reservoir modeling and simulation
\end{committeemember}

\begin{committeemember}{Anirban Manna}{MS from Indian Institute of Technology}{Wells Research Engineer at ExxonMobil}
Expertise includes FEA, fatigue, buckling, data analysis and drilling dynamics.
\end{committeemember}

\begin{committeemember}{Paul Pastusek}{MBA from University of Houston}{Drilling Mechanics Advisor at ExxonMobil}
Areas of interest include rig automation, debottlenecking, forensics analysis, drilling dynamics, borehole quality and technical training
\end{committeemember}

\begin{committeemember}{Greg Payette}{Ph.D. from Texas A\&M}{Wells Research Engineer at ExxonMobil}
Research interests include drilling dynamics, wellbore quality, open-source software, FEA and drilling automation
\end{committeemember}

\begin{committeemember}{Roman Shor}{Ph.D. from University of Texas}{Associate Professor at University of Calgary}
Research interests include in the areas of drill string dynamics modelling and control, drilling optimization and drilling systems automation.
\end{committeemember}

\section{Why Use Open-Source}
Most of the models reviewed as part of this project and both the models evaluated are open-source.  As such, it warrants answering the question, ``why \emph{use} open-source software?''  Or perhaps, more to the point for this industry, ``why \emph{invest} in open-source software?''

In general, open-source software offers numerous advantages.  It promotes transparency and collaboration, fosters collaboration and innovation, and is cost-effective.  As they relate to drilling dynamics software, these attributes make open-source software an attractive option.

\emph{Using} open-source software drilling dynamics models offers advantages over closed source.  With closed-source models you do not know how they work.  They are black boxes.  Undoubtedly, many of these models are very good.  But equally without doubt, they are not perfect.  All models have limitations; it is a fact of models.  And it is important to know what those limitations are to know when you can and when you cannot trust a model.  Open-source models provide transparency, and that transparency translates to trust.  Additionally, closed-source models cannot be modified, which prohibits collaboration and stifles innovation.

%They can also serve to interrogate closed-source models by comparing closed-source to open-source models.  In this way, they can lend some of their trust to the closed-source models.

Any \emph{investment} requires a justification of the value returned and open-source software is no exception.  For drilling dynamics software, that return comes from the collaboration with academia and members of industry.  This creates an increase in innovation and productivity as well as long term stability.

Collaboration among talented individuals can spark innovation and accelerate productivity.  When a diverse group of skilled minds come together, their unique perspectives and expertise can lead to creative solutions and breakthroughs that would be difficult to achieve in isolation.  Drilling dynamics experts are few and are dispersed across many organizations.  That creates a need for partnerships to maximize efficiency and create breakthroughs.  Industry programs such as \osdu{} demonstrate how open-source software and collaborative efforts can achieve more than a solitary entity could.

The boom-and-bust nature of the oil industry makes long term stability of any program difficult.  In down turns, projects are shut down, interrupting progress.  Employees are laid off.  Some of them leave the industry and never return; others land at another company but have to restart and redo work.  Open-source software fosters redundancy of developers by allowing multiple individuals or teams to contribute to a project without the risk of duplicating efforts.  With open access to the source code developers can collectively build upon existing work, share their enhancements, and avoid reinventing features and backslides in progress.

%The historical view of drilling dynamics models has placed an emphasis on trying to create differentiation.

On a final note, drilling mechanics and dynamics computer codes have been in use in the industry for over fifty years and some currently used commercial codes are 20 to 30 years old.  Despite this seemingly long history, they have failed to reach the predictive power of computer models used in industries such as aerospace or automotive.  Part of this is due to the difficulties in observing, measuring, and quantifying downhole parameters.  But a major cause is the fragmentation of programs.  This refers to both the isolated nature of programs and the interruptions cause by downturns.  Taken altogether, it seems clear that to achieve significant advances in the art, a new approach is required.

%is the isolated nature programs have worked in combined 

%But part of it results from lack of investment.  While companies understand that drilling dynamics are important, it doesn't seem 