\chapter{Model Review}
\label{ch:modelreview}
Below is a brief review of the models available on the \osdc{} Github site related to drilling dynamics.  In addition, some models were reviewed that were potentially available through industry-university research programs.  Some key features are noted as well as reasons for selecting or not selecting the model for further evaluation.  For this project, the key properties required to warrant selection for evaluation include:
\begin{bulletedlist}
	\item Be a drill string model (not a bit model)
	\item Include dynamics (not be a static model)
	\item Access to the source code is available
	\item Include the torsional degree of freedom
\end{bulletedlist}

\section{Wilson}
The basis for the model is the Ph.D. dissertation of Joshua Wilson (\referencename~\cite{ref:wilson2017a}).  The model was further developed at Scientific Drilling.  It is a full featured stiff-string model based on the finite element method.  It includes coupled flexibility of the drill string, geometric nonlinearity, wellbore contact, fluid mass, and damping from fluid, amongst other features.

The full featured model is not open-source.  Instead, a limited version has been released to the \osdc{}.  Specifically, it has the following limitations.
\begin{bulletedlist}
	\item Straight wellbores only
	\item Constant wellbore diameter
	\item No buckling
	\item No dynamic calculations
	\item Graphic user interfaces have been removed
	\item Advanced plotting features have been removed
\end{bulletedlist}
Specifically, since the dynamics have been removed, this model was not considered for evaluation.

\section{University of Texas Model}
A drill string code developed at the University of Texas is a six degree of freedom per node (axial, torsional and lateral) model (\referencename~\cite{ref:zhang2023a}) that states it is a stiff string model.  The paper documents the theory of the code, its usage and provides some information on expected behaviors.  The model seems promising, however, access to the model was not granted, therefore it could not be evaluated.

\section{Texas A\&M Model}
Texas A\&M developed a model to serve as simulator used for training horizontal directional drilling (\referencename~\cite{ref:losoya2020a}).  It was developed on top of a high-performance open-source library written in C++.  The model seems promising; it incorporates multibody physics, CFD and other solvers, and leverages GPU acceleration.  However, the full model has not been released.  Importantly for the purpose of this project, the dynamics portion has not been open-sourced.  Therefore, this model was not selected for evaluation.

\section{Tulsa University Models}
Two models from Tulsa University were evaluated.  The source code for these models is available through the TUDRP program. Ultimately, neither were selected for further evaluation.  The reviews of each and reasoning for not selecting them are presented below.

\subsection{Lateral Whirl Model}
The Ph.D. dissertation of Kriti Singh (\referencename~\cite{ref:singh2019a}) presents a model for lateral dynamics of drill  strings.  The model primarily is used to predict whirl in the drill string.  Dynamics are calculated in terms of Euler-Bernoulli beam elements and the solution is the general solution to the differential equation.  The Dynamic Transfer Matrix Simulator can be used to calculate the response of the BHA assembly to the specified excitation, for the specified $RPM$, WOB, and other operating parameters.  For the Dynamics Transfer Matrix Simulator it is assumed that if the deflection is larger than the borehole clearance, backward whirl will occur.  The numerical calculation is based on a single mass rotating disk.  This would only allow capturing the first natural frequency.  Since this model is a single degree of freedom lateral model, it does not fit the criteria of this project and it was not selected.

\subsection{Soft String Model}
\label{sec:tulasoftstring}
The soft string model from Tulsa University is based on \referencename~\cite{ref:miska2015a} where the model was originally developed.  This model is ``dynamic'' in the sense that it incorporates inertia effects.  This is an improvement on other models which are purely static.  A key assumption is that the pipe is not rotating, therefore, torsional oscillations are ignored.  Essentially, this becomes dynamics of a rigid body in the axial direction while accounting for the varied frictional effects (from string orientation) along the string.  Since this project was investigating torsional dynamic vibration models, this model was not suitable for evaluation.

In \referencename~\cite{ref:zamanipour2018a}, the model from \referencename~\cite{ref:miska2015a} is imporoved to add axial stiffness, static friction and drilling fluid drag.  The primary objective is calculation of forces for drill strings during tripping operations.  It uses a lumped mass and spring system.  It claims to be an axial ``stiff string'' model.  It allows for the string to be at the top or bottom of the hole automatically, however, it does not seem to account for the no contact case.  The wellbore is considered to be 2D.  A comparison is done between the axial ``stiff string'' model and a soft string model.  The improved model is an axially, multibody dynamic model.  However, torsional effects are still not considered.  Therefore, this model was not selected.

\input{"RGD Family of Models"}

\section{Aarsnes-Shor Models}
Aarsnes and Shor developed a distributed model for torsional dynamics of drill strings (\referencename~\cite{ref:aarsnes2017a}). This distributed model simulates the torsional dynamics by solving the angular velocity and torque governing equations using the finite difference method. Key assumptions of the model are the following.
\begin{bulletedlist}
	\item A soft string model (no lateral bending moments, entire string in contact with the borehole, lateral displacements are ignored)
	\item Axial displacements are ignored (Matlab version only)
	\item Bit-rock interaction is simplified to torque as a function of weight on bit
	\item The friction from the drill string and wellbore contact is modeled as Coulomb friction
    \item Viscous damping is included
    \item Cuttings are ignored
    \item Tool joint effects are not directly included and they must be manually accounted for
    \item Currently, only two drill pipe sizes of inner and outer diameter are allowed, one for the string and one for the BHA
    \item Cased and open hole sections have the same friction factor
\end{bulletedlist}

While the Coulomb friction model is somewhat less sophisticated than some models (e.g., a Stribeck model), reliable results were obtained from field validation (\referencename~\cite{ref:aarsnes2017a}).

Two versions of the source code are available, one implemented in Matlab and the other in Python.  Ongoing efforts are focused on enhancing the Python version by incorporating axial motion. Please see \chaptername~\ref{ch:aarnessshormodel} for a complete analysis.

\section{ExxonMobil Model}
The ExxonMobil model is based on a soft string axial-torsional model and utilizes a lumped mass and spring system (see \referencename~\cite{ref:dixit2021a}). It is designed to simulate the dynamics of a drill string under various drilling and tripping conditions for both vertical and inclined wells. It assumes the wellbore to be 2D and as such it is a 2-DOF per node system. The axial and torsional degrees of freedom are coupled through the friction terms. There is not a coupling from material properties or inertia terms.  The bit-rock interaction is modeled with depth of cut effects accounted for.  Key assumptions of the model are the following.
\begin{bulletedlist}
	\item A soft string model (no lateral bending moments, entire string in contact with the borehole, lateral displacements are ignored)
	\item The bit-rock interaction is simplified
	\item The friction from the drill string and wellbore contact is modeled as Stribeck friction
    \item Viscous damping is included
    \item Cuttings are ignored
    \item Cased and open hole sections have the same friction factor
\end{bulletedlist}

The governing equations are solved using a Runge-Kutta-Fehlberg method. The model has been implemented in Python.  Please see \chaptername~\ref{ch:exxonmobilmodel} for a complete analysis. 